Efficient High-Level Abstractions for Web Programming

### 2013-09-09 papers/gpce2013/gpce2013.tex
mention coffeescript, mention micro-benchmarks, update the figures

We validate our approach with a case study implemented with various candidate technologies and
discuss the relative pro and cons of them. Though the code written in our DSL is high-level and can
be shared between clients and servers, it has the same runtime performance on client-side as
hand-tuned low-level JavaScript code.

@@

The first approach for defining a cross-platform language consists in hard-coding, in the compiler,
the code generation scheme of each language feature to each target platform. Figure \ref{fig:langs}
(a) depicts this process. In order to support a feature related to a specific domain, the whole
compiler pipeline (parser, code generator, \etc) may have to be adapted. This approach gives
\emph{fat} languages because a lot of concepts are defined at the language level: general
programming concepts such as naming, functions, classes, as well as more domain-specific concepts
such as HTML fragment definition. Thus, implementing a fat language may require a high effort.
Examples of such languages for Web programming are Links~\cite{Cooper07_Links}, Opa and
Dart~\cite{Griffith11_Dart}.

@@

\code{null} references are a known source of problems in programming
languages~\cite{Hoare09_Null,Nanda09_Null}. For example, consider Listing \ref{null-unsafe} finding
a particular widget in the page and then a particular button within the widget. The native
\code{querySelector} method returns \code{null} if no node matched the given selector in the
document. If we run this code in a page where the widget is not present, it will throw an error
and stop further JavaScript execution. Defensive code can be written to handle \code{null}
references, but leads to very cumbersome code, as shown in Listing \ref{null-defensive}.

>>>>

We validate our approach with a case study implemented with various candidate technologies and
discuss the relative pro and cons of them. We also measured the individual impact of each of our optimizations using micro-benchmarks. Though the code written in our DSL is high-level and can
be shared between clients and servers, it has the same runtime performance on client-side as
hand-tuned low-level JavaScript code.

@@

The first approach for defining a cross-platform language consists in hard-coding, in the compiler,
the code generation scheme of each language feature to each target platform. Figure \ref{fig:langs}
(a) depicts this process. In order to support a feature related to a specific domain, the whole
compiler pipeline (parser, code generator, \etc) may have to be adapted. This approach gives
\emph{fat} languages because a lot of concepts are defined at the language level: general
programming concepts such as naming, functions, classes, as well as more domain-specific concepts
such as HTML fragment definition. Thus, implementing a fat language may require a high effort and adding support for these languages in development environments may require a even higher effort.
Examples of such languages for Web programming are Links~\cite{Cooper07_Links}, Opa, Dart~\cite{Griffith11_Dart}.

@@

\code{null} references are a known source of problems in programming
languages~\cite{Hoare09_Null,Nanda09_Null}. For example, consider Listing \ref{null-unsafe} finding
a particular widget in the page and then a particular button within the widget. The native
\code{querySelector} method returns \code{null} if no node matched the given selector in the
document. If we run this code in a page where the widget is not present, it will throw an error
and stop further JavaScript execution. Defensive code can be written to handle \code{null}
references, but leads to very cumbersome code, as shown in Listing \ref{null-defensive}.\footnote{However, one could alleviate the syntax burden by using a language such as CoffeeScript~\cite{Ashkenas2011_CoffeeScript}, that suppports a special notation for optional values dereferencing and desugars directly to JavaScript.}
